---
title: "Codebook"
author: "Shelbi"
date: "10/25/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Source

This project uses data from the Human Activity Recognition Using Smartphones Data Set (<http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones>). The data is included in the repo and can also be downloaded as a .zip here (<https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip>).


## About the Data Set

> The experiments have been carried out with a group of 30 volunteers within an age bracket of 19-48 years. Each person performed six activities (WALKING, WALKING_UPSTAIRS, WALKING_DOWNSTAIRS, SITTING, STANDING, LAYING) wearing a smartphone (Samsung Galaxy S II) on the waist. Using its embedded accelerometer and gyroscope, we captured 3-axial linear acceleration and 3-axial angular velocity at a constant rate of 50Hz. The experiments have been video-recorded to label the data manually. The obtained dataset has been randomly partitioned into two sets, where 70% of the volunteers was selected for generating the training data and 30% the test data. 

> The sensor signals (accelerometer and gyroscope) were pre-processed by applying noise filters and then sampled in fixed-width sliding windows of 2.56 sec and 50% overlap (128 readings/window). The sensor acceleration signal, which has gravitational and body motion components, was separated using a Butterworth low-pass filter into body acceleration and gravity. The gravitational force is assumed to have only low frequency components, therefore a filter with 0.3 Hz cutoff frequency was used. From each window, a vector of features was obtained by calculating variables from the time and frequency domain. See 'features_info.txt' for more details. 

## Libraries Used

dplyr is used to aggregate and conform variables into tidy data

```
library(dplyr)
```

## Data Used

run_analysis.R imports the subject (/test/subject_test.txt into ```SubjectTestData``` and /train/subject_train.txt into ```SubjectTrainData```), activity (/test/Y_test.txt into ```ActivityTestData``` and /train/Y_train.txt into ```ActivityTrainData```), and features (/test/X_test.txt into ```FeaturesTestData``` and /test/X_train.txt into ```FeaturesTrainData```) data accordingly into their own variables. The features labels (/features.txt) are read into ```FeaturesNamesData```.


# run_analysis.R

## Merge data into one table, extract only means and std measurements

Row data was merged using ```add_row``` from the ```dplyr``` package. 
Column names were assigned during this step using the chain function
These were stored in: ```ActivityData```, ```SubjectData```, ```FeatureData```


Data was then combined using ```col_bind``` from ```dplyr``` package
Data for only means and standard deviations for each feature were extracted in the same step by chaining the ```select``` function
This data was stored in: ```mergedData```


## Uses descriptive names for each activity
Activities, originally denoted by numbers, were changed into descriptive names.
"WALKING"
"WALKING_UPSTAIRS"
"WALKING_DOWNSTAIRS"
"SITTING"
"STANDING"
"LAYING"

## Appropriately labels data set with descriptive variable names
By examining mergedData, the following items were replaced using gsub:

```Acc``` replaced with Accelerometer
```Gyro``` replaced with Gyroscope
```BodyBody``` replaced with Body
```Mag``` replaced with Magnitude
Character ```f``` replaced with Frequency
Character ```t``` replaced with Time

variables after they were edited

```{r include=FALSE}
source("run_analysis.R")
```

```{r echo=TRUE}
names(mergedData)
```


## Create a second data set with the average of each variable for each activity and subject

Using ```dplyr``` functions ```group_by```, ```summarize_all``` and ```mean``` a new data set was created:

```
mean_Data
```


and written into the file "tidyData.txt"


